Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 interviews.tex
refine further

@@

But other users prefer to browse just the notebooks of the experts they
know. Rick says, ``usually I know somebody's notebooks that I want to search
through, like I'll search through Simon Urbanek's first, then I'll go off to
somebody else's, because I know that the kind of thing I'm looking for something
more obscure than I'm likely to find in some random person's.''

More selective ways to search will be needed as the volume of notebooks
continues to grow. We explore some ideas in the Discussion below.

\subsection{Integrated analysis}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull the
data, do the analytics on it, and then feed the viewer the data. So [...]
anyone from the stats group [...] can always insert something in between. We
were trying to make such an example so that they don't have to rewrite the GUI
part and they know how to get the data source. Once you get the data into
RCloud, then you have a dataframe to work with, and then they know how to
produce another dataframe.
\end{quote}

@@

However, the web interface of RCloud doesn't cut it for him:
\begin{quote}
I find that the style of editing is clunky. I would never edit like that if
I'm doing ordinary work. I usually use vi. [...] I mean, it's nice to have
it saved, but there's this trade-off between it making it easier for me to
present something or to save something, and my ease of typing and
correcting and things in a plain old editor window.
\end{quote}

Kenny also works with text file and command line, rearranging a file so the
good code is at the top of the file. Then he saves the file once he's got
the right code at the top, and often shares the work by pasting it into an
RCloud notebook.

@@

The interface takes some learning. Clint says
\begin{quote}
It took me a couple of weeks of
looking at it to become comfortable because I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing [...] I think I prefer it now, because you can cut your code up.
\end{quote}

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick explains:
\begin{quote}
If I was typing into one of the cells near that top, I had
to think ``I'm editing this cell and then I have to execute it,'' and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it. It was just another set of modes that I didn't like
\end{quote}

For some users, the cells getting saved feels incompatible with exploratory data
analysis. Rick notes that Markdown is a form of literate programming, and
``being literate about anything typically takes a lot of rewriting and going
back over things [...] I just want to explore a few things and then I'll know
what I want to write.''

Much of the time, commands typed at the R command line just serve a transitory
purpose, so having RCloud save commands into cells bothers experienced
users. Kenny notes, ``It saves everything I do like everything is gold, but most
of it is junk not meant to be saved.'' Rick says ``I just want to type in a
couple of quick commands and get some results that are going to tell
me what to do next, and they're not necessarily archival in any sense.''

The material that is not appropriate to save includes ``expressions that allow
me to check that I'm the right track'' (Rick), ``checking out what your data is,
or you make a plot of the data. Things that should not really become part of a
notebook, but things that help you understand your data better.'' (Raif)

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces that I'm thinking of on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell.''

Cells can also take up substantial vertical space, requiring scrolling. Rick
complains that ``every time I type more stuff, the notebook gets longer and
longer and it's harder to deal with.'' Ivan says that cells' controls and blank
space take extra real estate compared to a straight command line, and he and
KCKC both complained that when there are long results or plots, it causes the
code to scroll off the screen. Raif: ``for example imagine I run a cell whose
output is huge, and now that output would block the whole window, and to
find the next cell, I have to scroll down, and find where that cell starts,
right? So, I'm losing the continuity of my code.''

@@

To avert this danger, users are forced to write code within the notebook to
cache results. Kenny ends up ``littering the notebook with little switches that
comment out the slow parts'' and either ``load the object instead'' or ``save
the object to disk''. One of the authors of this paper manually writes cells
that check if a result file exists, and perform the calculation only if it's not
there.

@@

RCloud combines the functions of scratchpad with notebook assets. KC thinks
that having a scratch pad that doesn't persist between different notebooks is
``useless''.

@@

\subsection{A sea of notebooks}
RCloud is starting to be a victim of its own success, as there are now too many
notebooks.  There are over 5200 notebooks on the research instance, of which
more than 500 have been starred and more than 350 have been forked.

Rick doesn't find the notebook tree very useful for this reason: ``I
don't necessarily need to see everybody's notebook that uses RCloud [...] Every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.''

Although RCloud promises an environment where notebooks stay working, our users
have not all learned the habits that make this a reality. As Byers puts it,
there is still a big problem with ``bitrot'' - notebooks often stop working
because the user changed the structure of their data, changed a filename or a
database.  He says we need ``organizational protocols'' to catch up with the
technology.

There are technical problems to be solved here as well. Byers continues that the
problem with forking a notebook and correcting it, is that the original notebook
still exists with the error.  We need a way to see that there is a more recent
fork of notebook and to fetch in the changes from it.

Taylor, who writes packages and example notebooks for them, seems to be
haunted the most by dead notebooks.  At first, he had to learn how to write safe
notebooks: ``It would be a file inside a directory that only I had [...] access
to and so someone else couldn't run it, [...] a temporary file on something that
was read-only to everyone else. But I've gotten careful about that.''

But even correcting notebooks can come too late: ``I'll go back and change it and update
that page, but then what happens is people have in the meantime forked it. [...]
Usually what it'll be is that I'll have to change a package as well, so
their old fork stops working, and then they complain.''

Taylor tried to keep his notebook tree tidy, but this didn't help, because
people either keep old links in their email, or forked a notebook which is now
obsolete. By design, notebooks that are deleted in the user interface are not
purged from the repository. ``Now I'm really gun-shy about sharing notebooks
because it's like, do I want to support this forever?''

Taylor and Horace work together on a notebook:
\begin{quote}
There's no way for both of us
to have ownership of a notebook, so the only way is to fork it back and forth,
and so we have dozens of old copies. [We end up] deleting all the old ones, but
people still have links to them, because they don't actually disappear. [...]
The UI makes it easy to go back to the old stale versions, because you can say
`forked from' `forked from' `forked from', but there's actually no way to go the
way that we want people to go, which is up the tree.
\end{quote}

>>>>

Generally, analysts were more annoyed by the cell interface and the
perceived slowness of hte interface than web/tool developers. Analysts
switching to R from another language were more likely to use RCloud for
exploratory data analysis. Web/tool developers were more likely to see the
need for parameterization.

@@

But other users prefer to browse just the notebooks of the experts they
know. Rick says, ``usually I know somebody's notebooks that I want to search
through [...] because I know that the kind of thing I'm looking for is something
more obscure than I'm likely to find in some random person's.''

More selective ways to search will be needed as the volume of notebooks
continues to grow. We explore some ideas in Section~\ref{sec:discussion}.

\subsection{Integrated analysis}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  The structure of Horace's visualization
notebook means
\begin{quote}
the other guys who want to do analytics on the data, they can first pull
the data, do the analytics on it, and then feed the viewer the data. So
[...] anyone from the stats group [...] can always insert something in
between. [...] They don't have to rewrite the GUI part and they know how to
get the data source. Once you get the data into RCloud, then you have a
dataframe to work with, and then they know how to produce another
dataframe.
\end{quote}

@@

However, the web interface of RCloud doesn't work for him. ``It's nice to have
it saved, but there's this trade-off between it making it easier for me to
present something or to save something, and my ease of typing and
correcting and things in a plain old editor window.''

Kenny also works with text file and command line, rearranging a file so the
good code is at the top of the file. When he's done, all the the right is
code at the top, and often shares the work by pasting it into an RCloud
notebook.

@@

The interface takes some learning. Clint says:
\begin{quote}
It took me a couple of weeks of
looking at it to become comfortable because I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing.
\end{quote}

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick explains:
\begin{quote}
If I was typing into one of the cells near that top, I had
to think ``I'm editing this cell and then I have to execute it,'' and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it.
\end{quote}

For some users, the cells getting saved feels incompatible with exploratory data
analysis. Rick notes that Markdown is a form of literate programming, and
``being literate about anything typically takes a lot of rewriting and going
back over things. [...] I just want to explore a few things and then I'll know
what I want to write.''

Much of the time, commands typed at the R command line just serve a transitory
purpose, so having RCloud save commands into cells can be annoying. Kenny notes, ``It saves everything I do like everything is gold, but most
of it is junk not meant to be saved.'' Rick says ``I just want to type in a
couple of quick commands and get some results that are going to tell
me what to do next, and they're not necessarily archival in any sense.''

The material that is not appropriate to save includes ``expressions that allow
me to check that I'm the right track'' (Rick), ``checking out what your data is,
or you make a plot of the data. Things that should not really become part of a
notebook, but things that help you understand your data better'' (Raif).

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cell structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces [...] on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell.''

Cells can also take up substantial vertical space, requiring scrolling. Rick
complains that ``every time I type more stuff, the notebook gets longer and
longer and it's harder to deal with.'' Ivan says that cells' controls and blank
space take extra screen space compared to a straight command line, and he and
KC both complained that when there are long results or plots, it causes the
code to scroll off the screen. Raif explains, ``Imagine I run a cell whose
output is huge. [...] To find the next cell, I have to scroll down and find
where that cell starts. So I'm losing the continuity of my code.''

@@

To avert this danger, some users write code within the cells to
cache results. Kenny ends up ``littering the notebook with little switches that
comment out the slow parts'' and either ``load the object instead'' or ``save
the object to disk''. One of the authors of this paper manually writes cells
that check if a result file exists, and perform the calculation only if it's not
there.

@@

%% RCloud combines the functions of scratchpad with notebook assets. KC thinks
%% that having a scratch pad that doesn't persist between different notebooks is
%% ``useless''.

@@

\subsection{A sea of notebooks}
RCloud is starting to be a victim of its own success, as it is becoming
difficult to navigate all the notebooks.  There are over 5200 notebooks on
the research instance, of which more than 500 have been starred and more
than 350 have been forked.

For this reason, Rick doesn't find the notebook tree very useful:
\begin{quote}
I don't necessarily need to see everybody's notebook that uses RCloud [...] Every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.
\end{quote}

Although RCloud promises an environment where notebooks stay working, our
users have not all learned the habits that make this a reality. As Byers
puts it, there is still a big problem with ``bitrot'' -- notebooks often
stop working because the user changed the structure of their data, or
changed a filename or a database.  He says we need ``organizational
protocols'' to catch up with the technology. Even if one forks someone
else's notebook and corrects it, the original notebook still exists with
the error.

%% At first, he had to learn how to write safe
%% notebooks: ``It would be a file inside a directory that only I had [...] access
%% to and so someone else couldn't run it, [...] a temporary file on something that
%% was read-only to everyone else. But I've gotten careful about that.''

Taylor, who writes packages and example notebooks for them, seems to be
haunted the most by dead notebooks. Taylor and Horace work together on a notebook:
\begin{quote}
There's no way for both of us to have ownership of a notebook, so the only
way is to fork it back and forth, and so we have dozens of old copies. [We
end up] deleting all the old ones, but people still have links to them,
because they don't actually disappear.
\end{quote}

The problem of dead notebooks is compounded by changing
packages and their example notebooks at the same time. Taylor will ``go back and
change it and update the page, but then what happens is people have in the
meantime forked it. [...]  I'll have to change a package as well, so their
old fork stops working, and then they complain.''

Taylor tried to keep his notebook tree tidy, but this didn't help, because
people either kept old links in their email, or forked a notebook which is now
obsolete. By design, notebooks that are deleted in the user interface are not
purged from the repository. ``Now I'm really gun-shy about sharing notebooks
because it's like, do I want to support this forever?''
